Large-Scale Content-Based Matching of MIDI and Audio Files

### 2015-05-01 paper.tex
Fix formatting/typos

For our purposes, the reliability of the DTW confidence score is crucial because we will use it to decide when an audio/MIDI file pairing is valid and when an alignment can be used as training data for our hashing model.
We experimented with a variety of approaches, and converged on the following system for aligning a single MIDI/audio file pair:
First, we synthesize the MIDI data using \texttt{fluidsynth}\footnote{\texttt{http://www.fluidsynth.org}}.
We then estimate the MIDI beat locations using the MIDI file's tempo change information and the method described in \cite{raffel2014pretty_midi}.
To circumvent the common issue where the beat is tracked one-half beat out of phase, we double the BPM until it was at least 240.
Using \texttt{librosa} \cite{mcfee2014librosa}, we compute beat locations for the audio signal with the constraint that the BPM should remain close to the global MIDI tempo. 
We then compute log-amplitude beat-synchronous constant-Q transforms (CQTs) of audio and synthesized MIDI data with semitone frequency spacing and a frequency range from C3 (65.4 Hz) to C7 (1046.5 Hz), also using \texttt{librosa}.
The resulting feature matrices are then of dimensionality $N \times D$ and $M \times D$ where $N$ and $M$ are the resulting number of beats in the MIDI and audio recordings respectively and $D$ is $48$ (the number of semitones between C3 and C7).

@@

Unfortunately, the method described above does not meet this need.
The median number of beats in the 7digital preview clips from the MSD and the MIDI files are 186 and 1218 respectively.
Computing a cosine similarity matrix of this size (with a feature dimensionality of 48) using the highly optimized C++ code from \texttt{scipy} \cite{jones2014scipy} takes on average 9.82 milliseconds on a Intel Core i7-4930k processor.
When implemented using the LLVM just-in-time compiler Python module \texttt{numba}\footnote{\texttt{http://numba.pydata.org/}}, the DTW cost calculation described above takes on average 892 microseconds on the same processor.
It follows that matching a \textit{single} MIDI file to the MSD using this approach would take just under three hours.
Clearly, a more efficient approach is necessary.

@@

In order to apply this technique, we also need to obtain the collections of pairs similar and dissimilar feature vectors $\mathcal{P}$ and $\mathcal{N}$.
In our problem, our two feature modalities are CQTs from synthesized MIDI files and CQTs from audio files.
A logical choice for $\mathcal{P}$ is to choose pairs of constant-Q spectra from simultaneous beats in successfully aligned MIDI/audio pairs.
For example, if $A, B \in \mathbb{R}^{N \times D}$ are CQTs of an aligned MIDI and audio file respectively, then $\mathcal{P}$ will include $(A[1], B[1]), \ldots, (A[N], B[N])$.
The choice of $\mathcal{N}$ is less obvious.
We found that simply randomly choosing CQT spectra from non-aligned beats in our collection of aligned MIDI and audio files achieved satisfactory results.

@@

\begin{figure}[h!]
  \label{fig:percent_below_rank}
  \centering
    \includegraphics[width=0.5\textwidth]{percent_below_rank.png}
  \caption{Percentage of hash matches below rank}
\end{figure}

\begin{figure}[h!]
  \label{fig:rank_vs_score}
  \centering
    \includegraphics[width=0.5\textwidth]{rank_vs_score.png}
  \caption{Hash match rank vs. DTW score with color corresponding to correct tempo}
\end{figure}

>>>>

For our purposes, the reliability of the DTW confidence score is crucial because we will use it to decide when an audio/MIDI file pairing is valid and when an alignment can be used as training data for our hashing model.
We experimented with a variety of approaches, and converged on the following system for aligning a single MIDI/audio file pair:
First, we synthesize the MIDI data using \texttt{fluidsynth}.\footnote{\texttt{http://www.fluidsynth.org}}
We then estimate the MIDI beat locations using the MIDI file's tempo change information and the method described in \cite{raffel2014pretty_midi}.
To circumvent the common issue where the beat is tracked one-half beat out of phase, we double the BPM until it was at least 240.
Using \texttt{librosa} \cite{mcfee2014librosa}, we compute beat locations for the audio signal with the constraint that the BPM should remain close to the global MIDI tempo. 
We then compute log-amplitude beat-synchronous constant-Q transforms (CQTs) of audio and synthesized MIDI data with semitone frequency spacing and a frequency range from C3 (65.4 Hz) to C7 (1046.5 Hz), also using \texttt{librosa}.
The resulting feature matrices are then of dimensionality $N \times D$ and $M \times D$ where $N$ and $M$ are the resulting number of beats in the MIDI and audio recordings respectively and $D$ is $48$ (the number of semitones between C3 and C7).

@@

Unfortunately, the method described above does not meet this need.
The median number of beats in the 7digital preview clips from the MSD and the MIDI files are 186 and 1218 respectively.
Computing a cosine similarity matrix of this size (with a feature dimensionality of 48) using the highly optimized C++ code from \texttt{scipy} \cite{jones2014scipy} takes on average 9.82 milliseconds on a Intel Core i7-4930k processor.
When implemented using the LLVM just-in-time compiler Python module \texttt{numba},\footnote{\texttt{http://numba.pydata.org/}} the DTW cost calculation described above takes on average 892 microseconds on the same processor.
It follows that matching a \textit{single} MIDI file to the MSD using this approach would take just under three hours.
Clearly, a more efficient approach is necessary.

@@

In order to apply this technique, we also need to obtain the collections of pairs of similar and dissimilar feature vectors $\mathcal{P}$ and $\mathcal{N}$.
In our problem, our two feature modalities are CQTs from synthesized MIDI files and CQTs from audio files.
We therefore build $\mathcal{P}$ by choosing pairs of constant-Q spectra from simultaneous beats in successfully aligned MIDI/audio pairs.
For example, if $A, B \in \mathbb{R}^{N \times D}$ are CQTs of an aligned MIDI and audio file respectively, then $\mathcal{P}$ will include $(A[1], B[1]), \ldots, (A[N], B[N])$.
The choice of $\mathcal{N}$ is less obvious.
We found that simply randomly choosing CQT spectra from non-aligned beats in our collection of aligned MIDI and audio files achieved satisfactory results.

@@

\begin{figure}[h!]
  \label{fig:percent_below_rank}
  \centering
    \includegraphics[width=\columnwidth]{percent_below_rank.png}
  \caption{Percentage of hash matches below rank}
\end{figure}

\begin{figure}[h!]
  \label{fig:rank_vs_score}
  \centering
    \includegraphics[width=\columnwidth]{rank_vs_score.png}
  \caption{Hash match rank vs. DTW score with color corresponding to correct tempo}
\end{figure}
